---
title: "Customer Churn Rate"
author: "Valencia Lie"
date: "13 July 2020"
output:
  rmdformats::readthedown:
    highlight: kate
    toc: 6
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Business question of the day

Customer churn rate is the rate at which customers stop doing business with an entity. It is most commonly expressed as the percentage of service subscribers who discontinue their subscriptions within a given time period. It is vital for businesses to be able to predict whether their customers will churn or not, so that they will be able to improve their products or services if the predicted churn rate is high. They will also be able to cater to the customer's tastes and preferences more if they know the direct factors that affect the customer churn rate.

In this particular data set, the parameters are:

* CustomerID: Customer ID
* Gender: Whether the customer is a male or a female
* SeniorCitizen: Whether the customer is a senior citizen or not (1, 0)
* Partner: Whether the customer has a partner or not (Yes, No)
* Dependents: Whether the customer has dependents or not (Yes, No)
* Tenure: Number of months the customer has stayed with the company
* MultipleLines: Whether the customer has multiple lines or not (Yes, No, No phone service)
* OnlineSecurity: Whether the customer has online security or not 
* OnlineBackup: Whether the customer has online backup or not 
* DeviceProtection: Whether the customer has device protection or not 
* TechSupport: Whether the customer has tech support or not 
* StreamingTV: Whether the customer has streaming TV or not
* StreamingMovies: Whether the customer has streaming movies or not
* Contract: The contract term of the customer (Month-to-month, One year, Two year)
* PaperlessBilling: Whether the customer has paperless billing or not (Yes, No)
* PaymentMethod: The customer’s payment method (Electronic check, Mailed check, Bank transfer (automatic), Credit card (automatic))
* MonthlyCharges: The amount charged to the customer monthly
* TotalCharges: The total amount charged to the customer
* Churn: Whether the customer churned or not (Yes or No)

In the report below, I will attempt to predict whether a customer will churn or not based on some of these parameters using logistic regression and k-Nearest-Neighbour (kNN) model and will check how reliable my model is in predicting an outcome.

# Structure of this report
- Importing dataset and cleansing 
- Exploratory data analysis:
  - Proportion of data
  - Distribution of data analysis
- Cross Validation:
  - Recheck proportion of data
- Building a logistic regression model:
  - Insights of coefficients
  - Check for perfect separation
- Building a kNN model:
  - Pre-processing
  - Standardisation
  - Picking of optimal k
- Predicting future data:
  - Logistic Regression
- Evaluation:
  - For logistic regression model:
    1. Residuals
    2. 3 Assumptions
    3. Accuracy, Sensitivity and Precision (using confusion matrix)
  - For kNN:
    1. Accuracy, Sensitivity and Precision (using confusion matrix)
- Final model
- Model tuning
- Final model + conclusion

#Importing dataset and cleansing

```{r}
library(tidyverse)
churn <- read_csv("watson-churn.csv")
```

```{r}
head(churn)
```

```{r}
churn <- churn %>% 
  select(-customerID) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(SeniorCitizen = as.factor(SeniorCitizen))
head(churn)
```

```{r}
anyNA(churn)
churn %>% 
  is.na() %>% 
  colSums()

churn <- churn %>% 
  drop_na()
```

#Exploratory data analysis

##Proportion of data
```{r}
round(prop.table(table(churn$Churn)),2)
```

Even though the proportion is not perfect at 50:50, it is still considered balance as an imbalance data is usually 90:10 or 95:5.

##Distribution of data analysis
```{r}
summary(churn)
```

Since the numeric ranges between predictors vary greatly, hence we would need to process the data in order to standardise them into the same range using Z score standardisation to build a kNN model.

# Cross validation
```{r}
library(rsample)
set.seed(100)
idx <- initial_split(data = churn, strata = Churn , prop = 0.8)
test <- testing(idx)
train <- training(idx)
```

## Rechecking proportion of data
```{r}
round(prop.table(table(test$Churn)),2)
```

```{r}
round(prop.table(table(train$Churn)),2)
```

Since the proportion matched before cross-validation, we are ready to move on.

#Building a logistic regression model

Using stepwise regression,

##Backward

```{r}
model_all <- glm(formula = Churn ~., data = train, family = "binomial")
model_none <- glm(Churn ~ 1, data = train, family = "binomial")
```

```{r}
backward <- step(model_all, direction = "backward")
summary(backward)
```

According to this backward model, the AIC is 3677.15 and the model has a residual deviance of 3641.1. 
Let's try to compare it with a forward and both model.

## Forward
```{r}
forward <- step(model_none, direction = "forward", scope = list(lower = model_none, upper = model_all))
summary(forward)
```

This forward model has an AIC of 3677.15 and residual deviance of 3641.1.


##Both 
```{r}
both <- step(model_none, direction = "both", scope = list(lower = model_none, upper = model_all))
summary(both)
```

The both model has an AIC of 3677.15 and residual deviance of 3641.1. This shows that all backward, forward and both models have the same AIC and residual deviance. Hence, for simplicity sake, we will move forward with the backward model as our logistic regression model.

## Insights of coefficients
Since there are a lot of predictors being used, for simplicity sake, we will try to interpret only 3 predictors: namely TechSupport, Monthly charges and Contractyear.

```{r}
#Tech support
library(gtools)
exp(-0.5567838)
inv.logit(-0.5567838)

```
According to the above calculation, we can say that when a customer has tech support, he/she is 0.573 times more likely to churn, assuming that other variables are kept constant. This means that a customer is more likely to churn if he/she does NOT have tech support than if he/she does. We can also say that the probability of a customer churning if he/she has tech support is mere 0.364. Logically, this makes sense, because if a customer has plenty of tech support, he/she will be able to have a more enjoyable experience when using a certain product/service and thus will be less willing to churn.

```{r}
#Monthlycharges
exp(0.0261480)
```
According to the above calculation, an increase in $1 in the monthly charges charged to the customer, he/she is 1.026 more likely to churn, assuming that other variables are kept constant. Again, logically, this makes sense because the higher the subscription charges, customers will be less willing to continue using the product/services if they deem the price to be too expensive.

```{r}
#Contractyear
exp(-0.5077005)
exp(-1.0193470)
```
According to the above calculation, we can say that when a customer is under a 1 year contract, he/she is 0.602 times more likely to churn than if he/she is under a month-to-month contract, assuming that other variables are kept constant. In addition, when a customer is under a 2 year contract, he/she is 0.361 times more likely to churn than if he/she is under a month-to-month contract, assuming that other variables are kept constant. This means that a person is more likely to churn if he/she is on month-to-month contract, followed by 1 year contract, followed by 2 year contract. Logically, this makes perfect sense, because the longer the contract, the amount of money needed to pay upfront is larger, and hence customers will be less willing to terminate their contract/subscription before their contract ends. On the other hand, if the contract is short, customers will feel less burdened with the amount of payment made and thus may be willing to churn more than if the customer is on a yearly contract.

## Checking for perfect separation
According to the summary of the model, there seems to not be any perfect separation that had occurred. This is because, none of the coefficients differ very greatly with each other and the residual deviance is not even close to 0. Thus, it is safe to conclude that there is no perfect separation.

# Building kNN model

## Pre-processing
```{r}
test_x <- test %>% 
  select_if(is.numeric)

test_y <- test %>%
  select_if(is.factor)

train_x <- train %>% 
  select_if(is.numeric)

train_y <- train %>% 
  select_if(is.factor)

train_x 
train_y

```

## Standardisation
```{r}
train_x <-  scale(train_x)
test_x <- scale(test_x,
      center = attr(train_x, "scaled:center"),
      scale = attr(train_x, "scaled:scale"))
```

##Optimum K
```{r}
sqrt(nrow(train_x))
```

```{r}
library(class)
knnmodel <- knn(train = train_x, test = test_x, cl = train_y$Churn, k = 63)
```

#Predicting Future data

## Logistic Regression

```{r}
test$prob <- predict(backward, newdata = test, type = "response")
test$label <- ifelse(test = test$prob > 0.5, "Yes", "No")
test$label <- as.factor(test$label)
```

# Evaluation

## Logistic regression

### Residuals
According to the summary of the backward model, it has a residual deviance of 3641.1. Since there is no benchmark on what is considered to be a low residual deviance, it is hard to conclude whether this model is good enough based on this evaluation technique alone. Hence, we will try to see whether it fulfills the other aspects of being a good logistic regression model.

### Three Assumptions:

#### Independent Observations
According to the data it is safe to conclude that each observation is independent with each other because it is unlikely for a customer's data to be dependent on another customer's data. Hence, we say that this model fulfills this particular assumption.

However, in the event that it is very hard to pinpoint whether the observations are independent, we can conduct chi-square test between all the predictors (two at a time only).

Chi-square test hypothesis
H0: The two variables are independent.
H1: The two variables relate to each other.

#### Linearity of Predictor & Log of Odds
Since we have cleaned the data and changed some of the data types into their respective suitable data types, it can be said that this model has achieved linearity of predictor and log of odds. 

#### No multicollinearity
```{r}
#Done in Rstudio cloud because my old R does not support the package
#library(car)
#vif(backward)
```
                      GVIF Df GVIF^(1/(2*Df))
SeniorCitizen     1.103626  1        1.050536
Dependents        1.052844  1        1.026082
tenure           34.097157  1        5.839277
MultipleLines     1.433372  1        1.197235
OnlineSecurity    1.097261  1        1.047502
OnlineBackup      1.245625  1        1.116076
DeviceProtection  1.336968  1        1.156273
TechSupport       1.158529  1        1.076350
StreamingMovies   1.769173  1        1.330103
Contract          1.763532  2        1.152380
PaperlessBilling  1.069084  1        1.033965
PaymentMethod     1.231277  3        1.035284
MonthlyCharges    3.815317  1        1.953284
TotalCharges     42.319693  1        6.505359

Besides TotalCharges and tenure, the rest has VIF of less than 10, making them non-multicollinear to each other. 

### Confusion Matrix
```{r}
library(caret)
confusionMatrix(data = test$label, reference = test$Churn, positive = "Yes")
```
Based on the confusion matrix above, we can tell that the accuracy of the model is 
0.7298, the recall is 0.5615 and the precision is 0.5933. Based on our business problem, it is vital to prioritise 'recall' above the other metrics because it is crucial for the company to lower down the occasion in which they mistakenly predict that customers will not churn when in reality they will. 

## kNN
### Confusion Matrix
```{r}
confusionMatrix(data = knnmodel, reference = test_y$Churn, positive = "Yes")
```
This kNN model has an accuracy of 0.7029, recall of 0.3880 and precision of 0.5694. Similarly with the backward model, it is vital to prioritise 'recall' over the other metrics. 

# Comparison
According to the above confusion matrix, we can tell that the backward model generally performs better than the kNN model. The backward model has higher accuracy (0.7298 vs 0.7029), higher recall (0.5615 vs 0.3880) and higher precision (0.5933 vs 0.5694). In addition, since we mainly prioritise 'recall', the high discrepancies between the two recall values make it certain that we should choose the backward model over the kNN model.

Moreover, since we know that we should prioritise recall, we are able to further tune our backward model (unlike how we can't with kNN unless we change the value of k around) in order to achieve better recall.

# Tuning model
In order to achieve the assumption of no multicollinearity, we will try to make a model that excludes both Totalcharges and tenure.

```{r}
final_model <- glm(Churn ~ SeniorCitizen + Dependents + tenure+ MultipleLines + OnlineSecurity + OnlineBackup + DeviceProtection + TechSupport + StreamingMovies + Contract + PaperlessBilling + PaymentMethod +MonthlyCharges, data = train, family = "binomial")
```

Now, we will try to test the no multi-collinearity assumption again using vif.
```{r}
#vif(final_model)
```
                     GVIF Df GVIF^(1/(2*Df))
SeniorCitizen    1.104169  1        1.050795
Dependents       1.052093  1        1.025716
tenure           2.265190  1        1.505055
MultipleLines    1.404251  1        1.185011
OnlineSecurity   1.091441  1        1.044721
OnlineBackup     1.220036  1        1.104552
DeviceProtection 1.308876  1        1.144061
TechSupport      1.148543  1        1.071701
StreamingMovies  1.713827  1        1.309132
Contract         1.661308  2        1.135305
PaperlessBilling 1.066827  1        1.032873
PaymentMethod    1.213847  3        1.032826
MonthlyCharges   2.722947  1        1.650136

Now, we can see that all the VIFs of the predictors are below 10, which means this model fulfills this assumption.

We can also tune our backward model to achieve better recall value by decreasing the threshold probability when predicting future data. Previously, we used the default 0.5 and so we can lower it down to 0.45 or 0.4 for better recall result. This is because, by lowering the threshold probability, we are lowering the possibility of false negative as compared to before.

```{r}
library(caret)
test$prob0 <- predict(final_model, newdata = test, type = "response")
test$label0 <- ifelse(test = test$prob0 > 0.5, "Yes", "No")
test$label0 <- as.factor(test$label0)
confusionMatrix(data = test$label0, reference = test$Churn, positive = "Yes")
```

```{r}
test$label1 <- ifelse(test = test$prob0 > 0.45, "Yes", "No")
test$label1 <- as.factor(test$label1)
confusionMatrix(data = test$label1, reference = test$Churn, positive = "Yes")
```

We can see that by lowering the probability threshold, the recall value has improved (0.6215 vs 0.5552), although accuracy (0.7329 to 0.7288) and precision has deteriorated slightly (0.6007 to 0.5811). 

We will try to lower down the probability threshold again to 0.4 and see whether it is worth doing that. 

```{r}
test$label2 <- ifelse(test = test$prob0 > 0.4, "Yes", "No")
test$label2 <- as.factor(test$label2)
confusionMatrix(data = test$label2, reference = test$Churn, positive = "Yes")
```
When we lower the threshold even more, we can see a dip in both the accuracy (to 0.7267) and precision (to 0.5696), though there is improvement in the recall value 
(to 0.6845). Since we are not able to win in all metrics, I would say that it is okay to lower the threshold to 0.4 as the dip in accuracy and precision is not that much and that there is some substantial improvement in the recall metrics (which is what we needed in this particular case). 

# Final model and conclusion

In conclusion, the final model I would use to predict whether a customer will or will not churn is the tuned final model (with TotalCharges and tenure deleted from the model and threshold of 0.4). This is because, not only does it have a better performance, in terms of metrics, compared to the kNN model, it also satisfies what constitutes to be a good logistic regression model. It has a reasonable residual deviance value, it fulfills all three assumptions and delivers a decent performance in terms of the recall metric. 

However, with that being said, there is always a caveat in all kinds of models; they are imperfect and will be riddled with issues in one way or another. After all, predictions are tough to make on something as abstract and unpredictable as the future. 
